---
title: "TCR_Concordance_base"
author: "altreuter"
date: "07/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Base Code for TCR Concordance Analysis
Whenever TCRseq is performed by a CIMAC, reference samples distributed by MDAnderson will be sequenced as well.  These reference samples are then compared with reference samples received from other CIMAC TCRseq runs to look for concordance in the results.  The first such concordance analysis was performed in September of 2021.  This Rmd file is a subset of that code which can be used for future concordance analyses.

In order to run this analysis, previous reference samples received should be copied to a folder along with the new reference sample.  The variable "indir" would then be set to point to this folder.  (See first code block below.)

This code is taken from analysis performed 9/17/2021.  This code will run through all files in indir and create the following plots and tables:
1) Pairwise correlation plots for the top 50, top 100 and top 500 clones for all possible pairs of files.  Correlation values shown in dotplots, heatmaps and a summary table
2) Total clones in each file
3) Productive Frequency Histogram for all reference sample files


# Read in files, create top 50, top 100 and top 500 tables.
This code block:
 * Sets indir and outdir
 * Reads all files in the indir
 * Creates separate tables with top 50, top 100 and top 500 clones in each file.  (Top clones by productive frequency)

```{r,eval=TRUE}
library(tidyverse)
library(dplyr)
library(ggplot2)

#set input directory
indir<-"/Users/jen/Dropbox/TCR_Concordance/concord_data"
outdir<-"/Users/jen/Dropbox/TCR_Concordance/concord_output/2021_09_14/"

#create read table function that can accept a variable as a table name
read_tcr<-function(k){
  tcr_table<-read.table(k,header=T,sep="\t")
  tcr_table<-tcr_table %>%
    select(productive_frequency, cdr3_amino_acid, v_gene, d_gene, j_gene) %>%
    filter(!is.na(productive_frequency))
  return(tcr_table)
}

# get list of filenames from input directory
ff <- list.files(path=indir, full.names=TRUE)

# apply read_tcr function to list of files, add file names to each list member
myfilelist <- lapply(ff, read_tcr)
names(myfilelist)<-list.files(path=indir, full.names=FALSE)

#create separate file list with top 50
read_tcr_top<-function(s){
  tcr_table_50<-read.table(s,header=T,sep="\t")
  tcr_table_50<-tcr_table_50 %>%
    select(productive_frequency, cdr3_amino_acid, v_gene, d_gene, j_gene) %>%
    filter(!is.na(productive_frequency)) %>%
    slice_max(order_by = productive_frequency, n = 50, with_ties = TRUE)
  return(tcr_table_50)
}

#create separate file list with top 100
read_tcr_top100<-function(s){
  tcr_table_100<-read.table(s,header=T,sep="\t")
  tcr_table_100<-tcr_table_100 %>%
    select(productive_frequency, cdr3_amino_acid, v_gene, d_gene, j_gene) %>%
    filter(!is.na(productive_frequency)) %>%
    slice_max(order_by = productive_frequency, n = 100, with_ties = TRUE)
  return(tcr_table_100)
}

#create separate file list with top 500
read_tcr_top500<-function(s){
  tcr_table_500<-read.table(s,header=T,sep="\t")
  tcr_table_500<-tcr_table_500 %>%
    select(productive_frequency, cdr3_amino_acid, v_gene, d_gene, j_gene) %>%
    filter(!is.na(productive_frequency)) %>%
    slice_max(order_by = productive_frequency, n = 500, with_ties = TRUE)
  return(tcr_table_500)
}

myfilelist_top50<- lapply(ff,read_tcr_top)
names(myfilelist_top50)<-list.files(path=indir, full.names=FALSE)

myfilelist_top100<-lapply(ff,read_tcr_top100)
names(myfilelist_top100)<-list.files(path=indir, full.names=FALSE)

myfilelist_top500<-lapply(ff,read_tcr_top500)
names(myfilelist_top500)<-list.files(path=indir, full.names=FALSE)
```

# Correlation scatter plots, heatmaps and tables
This code block creates correlation scatter plots comparing productive frequencies for the top 50, 100 and 500 clones.  It also creates a heatmap and a summary table with the correlation values for all pairwise comparisons.  At the end it also summarizes the number of clones in the final merge.  This gives and indication if there is a significant mismatch in the number of matching clones between any two files.  (e.g. the top 50 clones in one file --the file listed in rows -- are contained within the top 53 of the paired file.) 

Note many of the variables all say "withd" because the original analysis was performed excluding the d genes and this if code including the d genes.  (See code from 9/17/2021 analysis...including the d genes did not significantly change results, so ultimately the following code which includes the d genes was used.)

```{r}
library(gridExtra)
library(reshape2)


#create inner_join function that include d gene
merge_tcrs_withd<-function(dd1, dd2){
  merged_files_d<-inner_join(dd1, dd2, by=c("v_gene" = "v_gene", "j_gene" ="j_gene", "cdr3_amino_acid"= "cdr3_amino_acid","d_gene"="d_gene"))
  return(merged_files_d)
}

#make empty lists for plots and correlations
myplots_withd<-list()
mycorrs_withd<-list()
mydims_withd<-list()

#run loop such that i=top 50 clone tsvs and h=full clone list
#e.g. myplots[[1]][[2]]= a comparison of the top50 clones of file 1 to
# all clones in file 2

for (i in 1:length(myfilelist)){
  #create plot and correlation list for file i within myplot_list
  myplots_withd[[i]]<-list()
  mycorrs_withd[[i]]<-list()
  mydims_withd[[i]]<-list()
  
  for (h in 1:length(myfilelist)){
    if (i==h) next
    tcr_i_h_withd<-merge_tcrs_withd(myfilelist_top50[[i]],myfilelist[[h]])
    
    p_withd<-ggplot(tcr_i_h_withd, aes(x=productive_frequency.x, y= productive_frequency.y)) +
      geom_point()+
      scale_x_log10() +
      scale_y_log10() +
      geom_abline(slope = 1, intercept = 0, linetype=2, color = "red") +
      theme_classic(base_size = 14) +
      ggtitle("correlation of productive frequency of top 50 clones")
  
    correlation_withd<-cor(tcr_i_h_withd$productive_frequency.x, tcr_i_h_withd$productive_frequency.y)
    dimension50_withd<-dim(tcr_i_h_withd)
    
    myplots_withd[[i]][[h]]<-p_withd
    mycorrs_withd[[i]][[h]]<-correlation_withd
    mydims_withd[[i]][[h]]<-dimension50_withd
    
  }
}

names(myplots_withd)<-list.files(path=indir, full.names=FALSE)
names(mydims_withd)<-list.files(path=indir, full.names=FALSE)

NameList<-list.files(path=indir, full.names=FALSE)

# function to display all correlations and plots.
# will also write to file
disp_plots_corr_withd<-function(file1,file2){
  cat(paste("Top 50 clones of ",NameList[file1],"compared to clones of ",NameList[file2],"\n"))
  cat(paste("Pearson Correlation = ",mycorrs_withd[[file1]][[file2]]))
  print(myplots_withd[[file1]][[file2]])
  #print plots to file
  png_file<-paste(outdir,"top50",NameList[file1],"_",NameList[file2],".png",sep="")
  png(file=png_file,width=600, height=350)
  print(myplots_withd[[file1]][[file2]])
  dev.off()
}


for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h) next
    disp_plots_corr_withd(i,h)
  }
}


#make matrix from mycorrs such that rows = top clones data and cols = all clones datas
mycorrs_mx_withd<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mycorrs_mx_withd[i,h]<-NA
  }else{
      mycorrs_mx_withd[i,h]<-mycorrs_withd[[i]][[h]]}
  }
}


rownames(mycorrs_mx_withd)<-NameList
colnames(mycorrs_mx_withd)<-NameList

#prepare correlation matrix for heatmap
melted_corr_withd <- melt(mycorrs_mx_withd)

#plot correlation heatmap
ggplot(data = melted_corr_withd, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
  ggtitle("Pearson correlations for top 50 clones") + 
  xlab("") +
  ylab("")

# Heatmap not very informative, print table as well
mycorrs_mx_withd %>% 
  DT::datatable()

##############################################
### Repeat analysis with top 100 clones  #####
##############################################

#make empty lists for plots and correlations
myplots100_withd<-list()
mycorrs100_withd<-list()
mydims100_withd<-list()
#run loop such that i=top 100 clone tsvs and h=full clone list
#e.g. myplots[[1]][[2]]= a comparison of the top100 clones of file 1 to
# all clones in file 2

for (i in 1:length(myfilelist)){
  #create plot and correlation list for file i within myplot_list
  myplots100_withd[[i]]<-list()
  mycorrs100_withd[[i]]<-list()
  mydims100_withd[[i]]<-list()
  
  for (h in 1:length(myfilelist)){
    if (i==h) next
    tcr100_i_h_withd<-merge_tcrs_withd(myfilelist_top100[[i]],myfilelist[[h]])
    
    p100_withd<-ggplot(tcr100_i_h_withd, aes(x=productive_frequency.x, y= productive_frequency.y)) +
      geom_point()+
      scale_x_log10() +
      scale_y_log10() +
      geom_abline(slope = 1, intercept = 0, linetype=2, color = "red") +
      theme_classic(base_size = 14) +
      ggtitle("correlation of productive frequency of top 100 clones")
  
    correlation100_withd<-cor(tcr100_i_h_withd$productive_frequency.x, tcr100_i_h_withd$productive_frequency.y)
    dimensions100_withd<-dim(tcr100_i_h_withd)
    
    myplots100_withd[[i]][[h]]<-p100_withd
    mycorrs100_withd[[i]][[h]]<-correlation100_withd
    mydims100_withd[[i]][[h]]<-dimensions100_withd
    
  }
}

names(myplots100_withd)<-list.files(path=indir, full.names=FALSE)
names(mydims100_withd)<-list.files(path=indir, full.names=FALSE)

# function to display all correlations and plots for top100.
# will also write to file
disp_plots_corr100_withd<-function(file1,file2){
  cat(paste("Top 100 clones of ",NameList[file1],"compared to clones of ",NameList[file2],"\n"))
  cat(paste("Pearson Correlation = ",mycorrs100_withd[[file1]][[file2]]))
  print(myplots100_withd[[file1]][[file2]])
  #print plots to file
  png_file<-paste(outdir,"top100",NameList[file1],"_",NameList[file2],".png",sep="")
  png(file=png_file,width=600, height=350)
  print(myplots100_withd[[file1]][[file2]])
  dev.off()
}


for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h) next
    disp_plots_corr100_withd(i,h)
  }
}

#make matrix from mycorrs such that rows = top clones data and cols = all clones datas
mycorrs100_mx_withd<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mycorrs100_mx_withd[i,h]<-NA
  }else{
      mycorrs100_mx_withd[i,h]<-mycorrs100_withd[[i]][[h]]}
  }
}


rownames(mycorrs100_mx_withd)<-NameList
colnames(mycorrs100_mx_withd)<-NameList

#prepare correlation matrix for heatmap
melted_corr100_withd <- melt(mycorrs100_mx_withd)

#plot correlation heatmap
ggplot(data = melted_corr100_withd, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
  ggtitle("Pearson correlations for top 100 clones") + 
  xlab("") +
  ylab("")

# Heatmap not very informative, print table as well
mycorrs100_mx_withd %>% 
  DT::datatable()

##############################################
### Repeat analysis with top 500 clones  #####
##############################################

#make empty lists for plots and correlations
myplots500_withd<-list()
mycorrs500_withd<-list()
mydims500_withd<-list()

#run loop such that i=top 500 clone tsvs and h=full clone list
#e.g. myplots[[1]][[2]]= a comparison of the top500 clones of file 1 to
# all clones in file 2

for (i in 1:length(myfilelist)){
  #create plot and correlation list for file i within myplot_list
  myplots500_withd[[i]]<-list()
  mycorrs500_withd[[i]]<-list()
  mydims500_withd[[i]]<-list()
  
  for (h in 1:length(myfilelist)){
    if (i==h) next
    tcr500_i_h_withd<-merge_tcrs_withd(myfilelist_top500[[i]],myfilelist[[h]])
    
    p500_withd<-ggplot(tcr500_i_h_withd, aes(x=productive_frequency.x, y= productive_frequency.y)) +
      geom_point()+
      scale_x_log10() +
      scale_y_log10() +
      geom_abline(slope = 1, intercept = 0, linetype=2, color = "red") +
      theme_classic(base_size = 14) +
      ggtitle("correlation of productive frequency of top 500 clones")
  
    correlation500_withd<-cor(tcr500_i_h_withd$productive_frequency.x, tcr500_i_h_withd$productive_frequency.y)
    dimensions500_withd<-dim(tcr500_i_h_withd)
    
    myplots500_withd[[i]][[h]]<-p500_withd
    mycorrs500_withd[[i]][[h]]<-correlation500_withd
    mydims500_withd[[i]][[h]]<-dimensions500_withd
    
  }
}

names(myplots500_withd)<-list.files(path=indir, full.names=FALSE)
names(mydims500_withd)<-list.files(path=indir, full.names=FALSE)

# function to display all correlations and plots for top100.
# will also write to file
disp_plots_corr500_withd<-function(file1,file2){
  cat(paste("Top 500 clones of ",NameList[file1],"compared to clones of ",NameList[file2],"\n"))
  cat(paste("Pearson Correlation = ",mycorrs500_withd[[file1]][[file2]]))
  print(myplots500_withd[[file1]][[file2]])
  #print plots to file
  png_file<-paste(outdir,"top500",NameList[file1],"_",NameList[file2],".png",sep="")
  png(file=png_file,width=600, height=350)
  print(myplots500_withd[[file1]][[file2]])
  dev.off()
}


for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h) next
    disp_plots_corr500_withd(i,h)
  }
}

#make matrix from mycorrs such that rows = top clones data and cols = all clones datas
mycorrs500_mx_withd<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mycorrs500_mx_withd[i,h]<-NA
  }else{
      mycorrs500_mx_withd[i,h]<-mycorrs500_withd[[i]][[h]]}
  }
}


rownames(mycorrs500_mx_withd)<-NameList
colnames(mycorrs500_mx_withd)<-NameList

#prepare correlation matrix for heatmap
melted_corr500_withd <- melt(mycorrs500_mx_withd)

#plot correlation heatmap
ggplot(data = melted_corr500_withd, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
  ggtitle("Pearson correlations for top 500 clones") + 
  xlab("") +
  ylab("")

# Heatmap not very informative, print table as well
mycorrs500_mx_withd %>% 
  DT::datatable()


######### print dimensions to evaluate if some clones are being lost with the inner join

#make matrix from mydims such that rows = top clones data and cols = all clones datas
# for top50 with d genes
mydims_mx_withd<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mydims_mx_withd[i,h]<-NA
  }else{
      mydims_mx_withd[i,h]<-mydims_withd[[i]][[h]][[1]]}
  }
}

rownames(mydims_mx_withd)<-NameList
colnames(mydims_mx_withd)<-NameList

#print matrix
print("Total number of clones after merge.  Rows = Top 50 clone file.")
mydims_mx_withd %>% 
  DT::datatable()


# for top100 with d genes
mydims100_mx_withd<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mydims100_mx_withd[i,h]<-NA
  }else{
      mydims100_mx_withd[i,h]<-mydims100_withd[[i]][[h]][[1]]}
  }
}

rownames(mydims100_mx_withd)<-NameList
colnames(mydims100_mx_withd)<-NameList

#print matrix
print("Total number of clones after merge.  Rows = Top 100 clone file.")
mydims100_mx_withd %>% 
  DT::datatable()

# for top500 with d genes
mydims500_mx_withd<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mydims500_mx_withd[i,h]<-NA
  }else{
      mydims500_mx_withd[i,h]<-mydims500_withd[[i]][[h]][[1]]}
  }
}


rownames(mydims500_mx_withd)<-NameList
colnames(mydims500_mx_withd)<-NameList

#print matrix
print("Total number of clones after merge.  Rows = Top 500 clone file.")
mydims500_mx_withd %>% 
  DT::datatable()
```

# Productive Frequency Summaries
This code block calculates the mean productive frequencies, fraction of clones with productive frequencies <0.001 and  < 0.01 for all files (all clones,top50, top100, top500).

Currently there is no output created for these calculations.  However, I am leaving the code in place in case this is useful.

```{r}
 #calculate mean productive frequencies for all files
top50_clone_freq_mean<-mean(prod_freq_matrix[,"top50_mean"])
top100_clone_freq_mean<-mean(prod_freq_matrix[,"top100_mean"])
top500_clone_freq_mean<-mean(prod_freq_matrix[,"top500_mean"])
all_clone_freq_mean<-mean(prod_freq_matrix[,"all_clones_mean"])

# calculate fraction of clones with freq <0.001
top50_freq_one_tenth<-sum(myfilelist_top50[[1]]$productive_frequency <0.001)/sum(myfilelist_top50[[1]]$productive_frequency>0)
top100_freq_one_tenth<-sum(myfilelist_top100[[1]]$productive_frequency <0.001)/sum(myfilelist_top100[[1]]$productive_frequency>0)
top500_freq_one_tenth<-sum(myfilelist_top500[[1]]$productive_frequency <0.001)/sum(myfilelist_top500[[1]]$productive_frequency>0)
all_freq_one_tenth<-sum(myfilelist[[1]]$productive_frequency <0.001)/sum(myfilelist[[1]]$productive_frequency>0)

# calculate fraction of clones with freq <0.01
top50_freq_one<-sum(myfilelist_top50[[1]]$productive_frequency<0.01)/sum(myfilelist_top50[[1]]$productive_frequency>0)
top100_freq_one<-sum(myfilelist_top100[[1]]$productive_frequency<0.01)/sum(myfilelist_top100[[1]]$productive_frequency>0)
top500_freq_one<-sum(myfilelist_top500[[1]]$productive_frequency<0.01)/sum(myfilelist_top500[[1]]$productive_frequency>0)
all_freq_one<-sum(myfilelist[[1]]$productive_frequency<0.01)/sum(myfilelist[[1]]$productive_frequency>0)


```

# Total Clones
This code block outputs the total clones in each original file.

```{r}
# make table of original file lengths

total_clones<-matrix(nrow=length(myfilelist),ncol=1)
for (i in 1:length(myfilelist)){
  total_clones[i,1]<-dim(myfilelist[[i]])[[1]]
}

rownames(total_clones)<- NameList
colnames(total_clones)<-c("Total Clones")

#print matrix
print("Total clones in each file")
total_clones %>% 
  DT::datatable()
```

# Make a histogram of the productive frequencies of each sample
```{r}

#create empty data frame
prod_freq_df<-data.frame()


# add productive frequencies to the dataframe
for (i in 1:length(myfilelist)){
  #make a dataframe from the ith productive frequency numbers
  temp_df<-data.frame(productive_frequency=myfilelist[[i]]$productive_frequency)
  #add info about where data came from
  temp_df$fileName <- names(myfilelist[i])
  # merge dataframes to make one large dataframe
  prod_freq_df<- rbind(prod_freq_df,temp_df)
}

ggplot(prod_freq_df, aes(productive_frequency, fill = fileName)) + 
   geom_histogram(alpha = 0.5, bins=100, position = 'identity') +
   scale_x_log10() +
   geom_abline(slope = 1, intercept = 0, linetype=2) +
   theme_classic(base_size = 14) 

ggplot(prod_freq_df, aes(productive_frequency, fill = fileName)) + 
   geom_histogram(alpha = 0.5, bins=100) +
   scale_x_log10() +
   geom_abline(slope = 1, intercept = 0, linetype=2) +
   theme_classic(base_size = 14) 

ggplot(prod_freq_df, aes(productive_frequency, fill = fileName)) + 
   geom_histogram(alpha = 0.5, bins=50, position = 'identity') +
   scale_x_log10() +
   geom_abline(slope = 1, intercept = 0, linetype=2) +
   theme_classic(base_size = 14) 

```

