---
title: "TCR_Concordance_2"
author: "altreuter"
date: "9/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code (and analysis) from Ming Tang 

(see "TCR_concordance.html"  from 9/10/2021 stored in TCR_Concordance Dropbox)

----------
```{r}
library(tidyverse)

tcr1<- read_tsv("/Users/jen/Dropbox/TCR_Concordance/concord_data/E4412_1A_TIL_CONTROL_DNA.tsv", guess_max = 1000000)
tcr2<- read_tsv("/Users/jen/Dropbox/TCR_Concordance/concord_data/E4412_1B_TIL_CONTROL.tsv", guess_max = 1000000)

dim(tcr1)

dim(tcr2)

colnames(tcr1)

tcr1<- tcr1 %>%
  select(productive_frequency, cdr3_amino_acid, v_gene, d_gene, j_gene) %>%
  filter(!is.na(productive_frequency)) %>%
  mutate(sample = "1A")

tcr2<- tcr2 %>%
  select(productive_frequency, cdr3_amino_acid, v_gene, d_gene, j_gene) %>%
  filter(!is.na(productive_frequency)) %>%
  mutate(sample = "1B")

dim(tcr1)

dim(tcr2)

tcr1_2<- inner_join(tcr1, tcr2, by=c("v_gene" = "v_gene", "j_gene" ="j_gene", "cdr3_amino_acid"= "cdr3_amino_acid"))

ggplot(tcr1_2, aes(x=productive_frequency.x, y= productive_frequency.y)) +
  geom_point()+
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(slope = 1, intercept = 0, linetype=2, color = "red") +
  theme_classic(base_size = 14) +
    ggtitle("correlation of productive frequency of all overlapping clones")
    
cor(tcr1_2$productive_frequency.x, tcr1_2$productive_frequency.y)
```


```{r}
## check the productive frequecy distribution. most of them are below 0.1%
ggplot(tcr1, aes(x= productive_frequency))+
  geom_histogram(bins=100) +
  scale_x_log10() +
  geom_abline(slope = 1, intercept = 0, linetype=2, color = "red") +
  theme_classic(base_size = 14)
```


```{r}
## 99% of the clones frequency < 0.01
mean(tcr1$productive_frequency <0.01)

## 82 clones have frequency > 0.001, so, checking top 50 or top 100 clones makes sense.
sum(tcr1$productive_frequency > 0.001)

## 76 clones
sum(tcr2$productive_frequency > 0.001)

### Check the correlation for the top 50 clones

## there are ties, so 53 clones were selected 
tcr1_top50<- tcr1 %>%
  slice_max(order_by = productive_frequency, n = 50, with_ties = TRUE)

tcr1_2_top50<- inner_join(tcr1_top50, tcr2, by=c("v_gene" = "v_gene", "j_gene" ="j_gene", "cdr3_amino_acid"= "cdr3_amino_acid"))

## 53 clones appear in both 1A and 1B samples!
dim(tcr1_2_top50)

ggplot(tcr1_2_top50, aes(x=productive_frequency.x, y= productive_frequency.y)) +
  geom_point()+
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(slope = 1, intercept = 0, linetype=2, color = "red") +
  theme_classic(base_size = 14) +
  ggtitle("correlation of productive frequency of top 50 clones")
  

cor(tcr1_2_top50$productive_frequency.x, tcr1_2_top50$productive_frequency.y)
```

# New Code (Attempt to make loop for all files in a directory)

```{r,eval=TRUE}
library(tidyverse)
library(dplyr)
library(ggplot2)

#set input directory
indir<-"/Users/jen/Dropbox/TCR_Concordance/concord_data"
outdir<-"/Users/jen/Dropbox/TCR_Concordance/concord_output/2021_09_14/"

#create read table function that can accept a variable as a table name
read_tcr<-function(k){
  tcr_table<-read.table(k,header=T,sep="\t")
  tcr_table<-tcr_table %>%
    select(productive_frequency, cdr3_amino_acid, v_gene, d_gene, j_gene) %>%
    filter(!is.na(productive_frequency))
  return(tcr_table)
}

# get list of filenames from input directory
ff <- list.files(path=indir, full.names=TRUE)

# apply read_tcr function to list of files, add file names to each list member
myfilelist <- lapply(ff, read_tcr)
names(myfilelist)<-list.files(path=indir, full.names=FALSE)

#create separate file list with top 50
read_tcr_top<-function(s){
  tcr_table_50<-read.table(s,header=T,sep="\t")
  tcr_table_50<-tcr_table_50 %>%
    select(productive_frequency, cdr3_amino_acid, v_gene, d_gene, j_gene) %>%
    filter(!is.na(productive_frequency)) %>%
    slice_max(order_by = productive_frequency, n = 50, with_ties = TRUE)
  return(tcr_table_50)
}

#create separate file list with top 100
read_tcr_top100<-function(s){
  tcr_table_100<-read.table(s,header=T,sep="\t")
  tcr_table_100<-tcr_table_100 %>%
    select(productive_frequency, cdr3_amino_acid, v_gene, d_gene, j_gene) %>%
    filter(!is.na(productive_frequency)) %>%
    slice_max(order_by = productive_frequency, n = 100, with_ties = TRUE)
  return(tcr_table_100)
}

#create separate file list with top 500
read_tcr_top500<-function(s){
  tcr_table_500<-read.table(s,header=T,sep="\t")
  tcr_table_500<-tcr_table_500 %>%
    select(productive_frequency, cdr3_amino_acid, v_gene, d_gene, j_gene) %>%
    filter(!is.na(productive_frequency)) %>%
    slice_max(order_by = productive_frequency, n = 500, with_ties = TRUE)
  return(tcr_table_500)
}

myfilelist_top50<- lapply(ff,read_tcr_top)
names(myfilelist_top50)<-list.files(path=indir, full.names=FALSE)

myfilelist_top100<-lapply(ff,read_tcr_top100)
names(myfilelist_top100)<-list.files(path=indir, full.names=FALSE)

myfilelist_top500<-lapply(ff,read_tcr_top500)
names(myfilelist_top500)<-list.files(path=indir, full.names=FALSE)
```

```{r, eval=FALSE, echo=FALSE}
#create inner_join function
merge_tcrs<-function(dd1, dd2){
  merged_files<-inner_join(dd1, dd2, by=c("v_gene" = "v_gene", "j_gene" ="j_gene", "cdr3_amino_acid"= "cdr3_amino_acid"))
  return(merged_files)
}

#make empty lists for plots and correlations
myplots<-list()
mycorrs<-list()
mydims<-list()

#run loop such that i=top 50 clone tsvs and h=full clone list
#e.g. myplots[[1]][[2]]= a comparison of the top50 clones of file 1 to
# all clones in file 2

for (i in 1:length(myfilelist)){
  #create plot and correlation list for file i within myplot_list
  myplots[[i]]<-list()
  mycorrs[[i]]<-list()
  mydims[[i]]<-list()
  
  for (h in 1:length(myfilelist)){
    if (i==h) next
    tcr_i_h<-merge_tcrs(myfilelist_top50[[i]],myfilelist[[h]])
    
    p<-ggplot(tcr_i_h, aes(x=productive_frequency.x, y= productive_frequency.y)) +
      geom_point()+
      scale_x_log10() +
      scale_y_log10() +
      geom_abline(slope = 1, intercept = 0, linetype=2, color = "red") +
      theme_classic(base_size = 14) +
      ggtitle("correlation of productive frequency of top 50 clones")
  
    correlation<-cor(tcr_i_h$productive_frequency.x, tcr_i_h$productive_frequency.y)
    dimension50<-dim(tcr_i_h)
    
    myplots[[i]][[h]]<-p
    mycorrs[[i]][[h]]<-correlation
    mydims[[i]][[h]]<-dimension50
    
  }
}

names(myplots)<-list.files(path=indir, full.names=FALSE)
names(mydims)<-list.files(path=indir, full.names=FALSE)

NameList<-list.files(path=indir, full.names=FALSE)

# function to display all correlations and plots.
# will also write to file
disp_plots_corr<-function(file1,file2){
  cat(paste("Top 50 clones of ",NameList[file1],"compared to clones of ",NameList[file2],"\n"))
  cat(paste("Pearson Correlation = ",mycorrs[[file1]][[file2]]))
  print(myplots[[file1]][[file2]])
  #print plots to file
  png_file<-paste(outdir,"top50",NameList[file1],"_",NameList[file2],".png",sep="")
  png(file=png_file,width=600, height=350)
  print(myplots[[file1]][[file2]])
  dev.off()
}


for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h) next
    disp_plots_corr(i,h)
  }
}

#make matrix from mycorrs such that rows = top clones data and cols = all clones datas
mycorrs_mx<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mycorrs_mx[i,h]<-NA
  }else{
      mycorrs_mx[i,h]<-mycorrs[[i]][[h]]}
  }
}


rownames(mycorrs_mx)<-NameList
colnames(mycorrs_mx)<-NameList

#prepare correlation matrix for heatmap

library(gridExtra)
library(reshape2)
melted_corr <- melt(mycorrs_mx)

#plot correlation heatmap
ggplot(data = melted_corr, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
  ggtitle("Pearson correlations for top 50 clones") + 
  xlab("") +
  ylab("")

# Heatmap not very informative, maybe print table?
mycorrs_mx %>% 
  DT::datatable()

##############################################
### Repeat analysis with top 100 clones  #####
##############################################

#make empty lists for plots and correlations
myplots100<-list()
mycorrs100<-list()
mydims100<-list()
#run loop such that i=top 100 clone tsvs and h=full clone list
#e.g. myplots[[1]][[2]]= a comparison of the top100 clones of file 1 to
# all clones in file 2

for (i in 1:length(myfilelist)){
  #create plot and correlation list for file i within myplot_list
  myplots100[[i]]<-list()
  mycorrs100[[i]]<-list()
    mydims100[[i]]<-list()
  
  for (h in 1:length(myfilelist)){
    if (i==h) next
    tcr100_i_h<-merge_tcrs(myfilelist_top100[[i]],myfilelist[[h]])
    
    p100<-ggplot(tcr100_i_h, aes(x=productive_frequency.x, y= productive_frequency.y)) +
      geom_point()+
      scale_x_log10() +
      scale_y_log10() +
      geom_abline(slope = 1, intercept = 0, linetype=2, color = "red") +
      theme_classic(base_size = 14) +
      ggtitle("correlation of productive frequency of top 100 clones")
  
    correlation100<-cor(tcr100_i_h$productive_frequency.x, tcr100_i_h$productive_frequency.y)
    dimensions100<-dim(tcr100_i_h)
    
    myplots100[[i]][[h]]<-p100
    mycorrs100[[i]][[h]]<-correlation100
    mydims100[[i]][[h]]<-dimensions100
    
  }
}

names(myplots100)<-list.files(path=indir, full.names=FALSE)
names(mydims100)<-list.files(path=indir, full.names=FALSE)

# function to display all correlations and plots for top100.
# will also write to file
disp_plots_corr100<-function(file1,file2){
  cat(paste("Top 100 clones of ",NameList[file1],"compared to clones of ",NameList[file2],"\n"))
  cat(paste("Pearson Correlation = ",mycorrs100[[file1]][[file2]]))
  print(myplots100[[file1]][[file2]])
  #print plots to file
  png_file<-paste(outdir,"top100",NameList[file1],"_",NameList[file2],".png",sep="")
  png(file=png_file,width=600, height=350)
  print(myplots100[[file1]][[file2]])
  dev.off()
}


for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h) next
    disp_plots_corr100(i,h)
  }
}

#make matrix from mycorrs such that rows = top clones data and cols = all clones datas
mycorrs100_mx<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mycorrs100_mx[i,h]<-NA
  }else{
      mycorrs100_mx[i,h]<-mycorrs100[[i]][[h]]}
  }
}


rownames(mycorrs100_mx)<-NameList
colnames(mycorrs100_mx)<-NameList

#prepare correlation matrix for heatmap
melted_corr100 <- melt(mycorrs100_mx)

#plot correlation heatmap
ggplot(data = melted_corr100, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
  ggtitle("Pearson correlations for top 100 clones") + 
  xlab("") +
  ylab("")

# Heatmap not very informative, maybe print table?
mycorrs100_mx %>% 
  DT::datatable()

##############################################
### Repeat analysis with top 500 clones  #####
##############################################

#make empty lists for plots and correlations
myplots500<-list()
mycorrs500<-list()
mydims500<-list()

#run loop such that i=top 500 clone tsvs and h=full clone list
#e.g. myplots[[1]][[2]]= a comparison of the top500 clones of file 1 to
# all clones in file 2

for (i in 1:length(myfilelist)){
  #create plot and correlation list for file i within myplot_list
  myplots500[[i]]<-list()
  mycorrs500[[i]]<-list()
  mydims500[[i]]<-list()
  
  for (h in 1:length(myfilelist)){
    if (i==h) next
    tcr500_i_h<-merge_tcrs(myfilelist_top500[[i]],myfilelist[[h]])
    
    p500<-ggplot(tcr500_i_h, aes(x=productive_frequency.x, y= productive_frequency.y)) +
      geom_point()+
      scale_x_log10() +
      scale_y_log10() +
      geom_abline(slope = 1, intercept = 0, linetype=2, color = "red") +
      theme_classic(base_size = 14) +
      ggtitle("correlation of productive frequency of top 500 clones")
  
    correlation500<-cor(tcr500_i_h$productive_frequency.x, tcr500_i_h$productive_frequency.y)
    dimensions500<-dim(tcr500_i_h)
    
    myplots500[[i]][[h]]<-p500
    mycorrs500[[i]][[h]]<-correlation500
    mydims500[[i]][[h]]<-dimensions500
    
  }
}

names(myplots500)<-list.files(path=indir, full.names=FALSE)
names(mydims500)<-list.files(path=indir, full.names=FALSE)

# function to display all correlations and plots for top100.
# will also write to file
disp_plots_corr500<-function(file1,file2){
  cat(paste("Top 500 clones of ",NameList[file1],"compared to clones of ",NameList[file2],"\n"))
  cat(paste("Pearson Correlation = ",mycorrs500[[file1]][[file2]]))
  print(myplots500[[file1]][[file2]])
  #print plots to file
  png_file<-paste(outdir,"top500",NameList[file1],"_",NameList[file2],".png",sep="")
  png(file=png_file,width=600, height=350)
  print(myplots500[[file1]][[file2]])
  dev.off()
}


for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h) next
    disp_plots_corr500(i,h)
  }
}

#make matrix from mycorrs such that rows = top clones data and cols = all clones datas
mycorrs500_mx<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mycorrs500_mx[i,h]<-NA
  }else{
      mycorrs500_mx[i,h]<-mycorrs500[[i]][[h]]}
  }
}


rownames(mycorrs500_mx)<-NameList
colnames(mycorrs500_mx)<-NameList

#prepare correlation matrix for heatmap
melted_corr500 <- melt(mycorrs500_mx)

#plot correlation heatmap
ggplot(data = melted_corr500, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
  ggtitle("Pearson correlations for top 500 clones") + 
  xlab("") +
  ylab("")

# Heatmap not very informative, maybe print table?
mycorrs500_mx %>% 
  DT::datatable()
```


```{r}
########################################
##### Productive Frequency Ranges ######
########################################

NameList<-list.files(path=indir, full.names=FALSE)

# get ranges of productive frequencies
prod_freq_matrix<-matrix(nrow=length(myfilelist),ncol=12)
rownames(prod_freq_matrix)<-NameList
colnames(prod_freq_matrix)<-c("all_clones_min","all_clone_max","all_clones_mean","top500_min","top500_max","top500_mean","top100_min","top100_max","top100_mean", "top50_min","top50_max","top50_mean")

## range for all clones
for (i in 1:length(myfilelist)){
  prod_freq_matrix[i,1:3]<-c(range(myfilelist[[i]]$productive_frequency),mean(myfilelist[[i]]$productive_frequency))
  prod_freq_matrix[i,4:6]<-c(range(myfilelist_top500[[i]]$productive_frequency),mean(myfilelist_top500[[i]]$productive_frequency))
  prod_freq_matrix[i,7:9]<-c(range(myfilelist_top100[[i]]$productive_frequency),mean(myfilelist_top100[[i]]$productive_frequency))
  prod_freq_matrix[i,10:12]<-c(range(myfilelist_top50[[i]]$productive_frequency),mean(myfilelist_top50[[i]]$productive_frequency))
}

# print productive frequency summary table
prod_freq_matrix %>% 
  DT::datatable()

```

## Check Analysis using "d_gene" in addition to other matches.

In the initial analysis, it seemed as if including the d_gene in the inner_join  resulted in a lot of lost matches.  Double checking...

```{r}
library(gridExtra)
library(reshape2)


#create inner_join function that include d gene
merge_tcrs_withd<-function(dd1, dd2){
  merged_files_d<-inner_join(dd1, dd2, by=c("v_gene" = "v_gene", "j_gene" ="j_gene", "cdr3_amino_acid"= "cdr3_amino_acid","d_gene"="d_gene"))
  return(merged_files_d)
}

#make empty lists for plots and correlations
myplots_withd<-list()
mycorrs_withd<-list()
mydims_withd<-list()

#run loop such that i=top 50 clone tsvs and h=full clone list
#e.g. myplots[[1]][[2]]= a comparison of the top50 clones of file 1 to
# all clones in file 2

for (i in 1:length(myfilelist)){
  #create plot and correlation list for file i within myplot_list
  myplots_withd[[i]]<-list()
  mycorrs_withd[[i]]<-list()
  mydims_withd[[i]]<-list()
  
  for (h in 1:length(myfilelist)){
    if (i==h) next
    tcr_i_h_withd<-merge_tcrs_withd(myfilelist_top50[[i]],myfilelist[[h]])
    
    p_withd<-ggplot(tcr_i_h_withd, aes(x=productive_frequency.x, y= productive_frequency.y)) +
      geom_point()+
      scale_x_log10() +
      scale_y_log10() +
      geom_abline(slope = 1, intercept = 0, linetype=2, color = "red") +
      theme_classic(base_size = 14) +
      ggtitle("correlation of productive frequency of top 50 clones with d genes")
  
    correlation_withd<-cor(tcr_i_h_withd$productive_frequency.x, tcr_i_h_withd$productive_frequency.y)
    dimension50_withd<-dim(tcr_i_h_withd)
    
    myplots_withd[[i]][[h]]<-p_withd
    mycorrs_withd[[i]][[h]]<-correlation_withd
    mydims_withd[[i]][[h]]<-dimension50_withd
    
  }
}

names(myplots_withd)<-list.files(path=indir, full.names=FALSE)
names(mydims_withd)<-list.files(path=indir, full.names=FALSE)

# function to display all correlations and plots.
# will also write to file
disp_plots_corr_withd<-function(file1,file2){
  cat(paste("Top 50 clones of ",NameList[file1],"compared to clones of ",NameList[file2],"with d genes","\n"))
  cat(paste("Pearson Correlation = ",mycorrs_withd[[file1]][[file2]]))
  print(myplots_withd[[file1]][[file2]])
  #print plots to file
  png_file<-paste(outdir,"top50",NameList[file1],"_",NameList[file2],"withD.png",sep="")
  png(file=png_file,width=600, height=350)
  print(myplots_withd[[file1]][[file2]])
  dev.off()
}


for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h) next
    disp_plots_corr_withd(i,h)
  }
}


#make matrix from mycorrs such that rows = top clones data and cols = all clones datas
mycorrs_mx_withd<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mycorrs_mx_withd[i,h]<-NA
  }else{
      mycorrs_mx_withd[i,h]<-mycorrs_withd[[i]][[h]]}
  }
}


rownames(mycorrs_mx_withd)<-NameList
colnames(mycorrs_mx_withd)<-NameList

#prepare correlation matrix for heatmap
melted_corr_withd <- melt(mycorrs_mx_withd)

#plot correlation heatmap
ggplot(data = melted_corr_withd, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
  ggtitle("Pearson correlations for top 50 clones, d genes included") + 
  xlab("") +
  ylab("")

# Heatmap not very informative, maybe print table?
mycorrs_mx_withd %>% 
  DT::datatable()

##############################################
### Repeat analysis with top 100 clones  #####
##############################################

#make empty lists for plots and correlations
myplots100_withd<-list()
mycorrs100_withd<-list()
mydims100_withd<-list()
#run loop such that i=top 100 clone tsvs and h=full clone list
#e.g. myplots[[1]][[2]]= a comparison of the top100 clones of file 1 to
# all clones in file 2

for (i in 1:length(myfilelist)){
  #create plot and correlation list for file i within myplot_list
  myplots100_withd[[i]]<-list()
  mycorrs100_withd[[i]]<-list()
  mydims100_withd[[i]]<-list()
  
  for (h in 1:length(myfilelist)){
    if (i==h) next
    tcr100_i_h_withd<-merge_tcrs_withd(myfilelist_top100[[i]],myfilelist[[h]])
    
    p100_withd<-ggplot(tcr100_i_h_withd, aes(x=productive_frequency.x, y= productive_frequency.y)) +
      geom_point()+
      scale_x_log10() +
      scale_y_log10() +
      geom_abline(slope = 1, intercept = 0, linetype=2, color = "red") +
      theme_classic(base_size = 14) +
      ggtitle("correlation of productive frequency of top 100 clones with d genes")
  
    correlation100_withd<-cor(tcr100_i_h_withd$productive_frequency.x, tcr100_i_h_withd$productive_frequency.y)
    dimensions100_withd<-dim(tcr100_i_h_withd)
    
    myplots100_withd[[i]][[h]]<-p100_withd
    mycorrs100_withd[[i]][[h]]<-correlation100_withd
    mydims100_withd[[i]][[h]]<-dimensions100_withd
    
  }
}

names(myplots100_withd)<-list.files(path=indir, full.names=FALSE)
names(mydims100_withd)<-list.files(path=indir, full.names=FALSE)

# function to display all correlations and plots for top100.
# will also write to file
disp_plots_corr100_withd<-function(file1,file2){
  cat(paste("Top 100 clones of ",NameList[file1],"compared to clones of ",NameList[file2],"with d genes","\n"))
  cat(paste("Pearson Correlation = ",mycorrs100_withd[[file1]][[file2]]))
  print(myplots100_withd[[file1]][[file2]])
  #print plots to file
  png_file<-paste(outdir,"top100",NameList[file1],"_",NameList[file2],"withD.png",sep="")
  png(file=png_file,width=600, height=350)
  print(myplots100_withd[[file1]][[file2]])
  dev.off()
}


for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h) next
    disp_plots_corr100_withd(i,h)
  }
}

#make matrix from mycorrs such that rows = top clones data and cols = all clones datas
mycorrs100_mx_withd<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mycorrs100_mx_withd[i,h]<-NA
  }else{
      mycorrs100_mx_withd[i,h]<-mycorrs100_withd[[i]][[h]]}
  }
}


rownames(mycorrs100_mx_withd)<-NameList
colnames(mycorrs100_mx_withd)<-NameList

#prepare correlation matrix for heatmap
melted_corr100_withd <- melt(mycorrs100_mx_withd)

#plot correlation heatmap
ggplot(data = melted_corr100_withd, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
  ggtitle("Pearson correlations for top 100 clones with d genes") + 
  xlab("") +
  ylab("")

# Heatmap not very informative, maybe print table?
mycorrs100_mx_withd %>% 
  DT::datatable()

##############################################
### Repeat analysis with top 500 clones  #####
##############################################

#make empty lists for plots and correlations
myplots500_withd<-list()
mycorrs500_withd<-list()
mydims500_withd<-list()

#run loop such that i=top 500 clone tsvs and h=full clone list
#e.g. myplots[[1]][[2]]= a comparison of the top500 clones of file 1 to
# all clones in file 2

for (i in 1:length(myfilelist)){
  #create plot and correlation list for file i within myplot_list
  myplots500_withd[[i]]<-list()
  mycorrs500_withd[[i]]<-list()
  mydims500_withd[[i]]<-list()
  
  for (h in 1:length(myfilelist)){
    if (i==h) next
    tcr500_i_h_withd<-merge_tcrs_withd(myfilelist_top500[[i]],myfilelist[[h]])
    
    p500_withd<-ggplot(tcr500_i_h_withd, aes(x=productive_frequency.x, y= productive_frequency.y)) +
      geom_point()+
      scale_x_log10() +
      scale_y_log10() +
      geom_abline(slope = 1, intercept = 0, linetype=2, color = "red") +
      theme_classic(base_size = 14) +
      ggtitle("correlation of productive frequency of top 500 clones with d genes")
  
    correlation500_withd<-cor(tcr500_i_h_withd$productive_frequency.x, tcr500_i_h_withd$productive_frequency.y)
    dimensions500_withd<-dim(tcr500_i_h_withd)
    
    myplots500_withd[[i]][[h]]<-p500_withd
    mycorrs500_withd[[i]][[h]]<-correlation500_withd
    mydims500_withd[[i]][[h]]<-dimensions500_withd
    
  }
}

names(myplots500_withd)<-list.files(path=indir, full.names=FALSE)
names(mydims500_withd)<-list.files(path=indir, full.names=FALSE)

# function to display all correlations and plots for top100.
# will also write to file
disp_plots_corr500_withd<-function(file1,file2){
  cat(paste("Top 500 clones of ",NameList[file1],"compared to clones of ",NameList[file2],"with d genes","\n"))
  cat(paste("Pearson Correlation = ",mycorrs500_withd[[file1]][[file2]]))
  print(myplots500_withd[[file1]][[file2]])
  #print plots to file
  png_file<-paste(outdir,"top500",NameList[file1],"_",NameList[file2],"withD.png",sep="")
  png(file=png_file,width=600, height=350)
  print(myplots500_withd[[file1]][[file2]])
  dev.off()
}


for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h) next
    disp_plots_corr500_withd(i,h)
  }
}

#make matrix from mycorrs such that rows = top clones data and cols = all clones datas
mycorrs500_mx_withd<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mycorrs500_mx_withd[i,h]<-NA
  }else{
      mycorrs500_mx_withd[i,h]<-mycorrs500_withd[[i]][[h]]}
  }
}


rownames(mycorrs500_mx_withd)<-NameList
colnames(mycorrs500_mx_withd)<-NameList

#prepare correlation matrix for heatmap
melted_corr500_withd <- melt(mycorrs500_mx_withd)

#plot correlation heatmap
ggplot(data = melted_corr500_withd, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
  ggtitle("Pearson correlations for top 500 clones with d genes") + 
  xlab("") +
  ylab("")

# Heatmap not very informative, maybe print table?
mycorrs500_mx_withd %>% 
  DT::datatable()


######### print dimensions to evaluate if some clones are being lost with the inner join

#make matrix from mydims such that rows = top clones data and cols = all clones datas
# for top50 with d genes
mydims_mx_withd<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mydims_mx_withd[i,h]<-NA
  }else{
      mydims_mx_withd[i,h]<-mydims_withd[[i]][[h]][[1]]}
  }
}

rownames(mydims_mx_withd)<-NameList
colnames(mydims_mx_withd)<-NameList

#print matrix
print("Total number of clones after merge.  Rows = Top 50 clone file.")
mydims_mx_withd %>% 
  DT::datatable()


# for top100 with d genes
mydims100_mx_withd<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mydims100_mx_withd[i,h]<-NA
  }else{
      mydims100_mx_withd[i,h]<-mydims100_withd[[i]][[h]][[1]]}
  }
}

rownames(mydims100_mx_withd)<-NameList
colnames(mydims100_mx_withd)<-NameList

#print matrix
print("Total number of clones after merge.  Rows = Top 100 clone file.")
mydims100_mx_withd %>% 
  DT::datatable()

# for top500 with d genes
mydims500_mx_withd<-matrix(nrow=length(myfilelist),ncol=length(myfilelist))
for (i in 1:length(myfilelist)){
  for (h in 1:length(myfilelist)){
    if (i==h){
      mydims500_mx_withd[i,h]<-NA
  }else{
      mydims500_mx_withd[i,h]<-mydims500_withd[[i]][[h]][[1]]}
  }
}


rownames(mydims500_mx_withd)<-NameList
colnames(mydims500_mx_withd)<-NameList

#print matrix
print("Total number of clones after merge.  Rows = Top 500 clone file.")
mydims500_mx_withd %>% 
  DT::datatable()
```

Including D genes does not seem significantly affect the results.  Marking earlier code as eval=FALSE.  Will knit with just with d genes code evaluated.

```{r}
 #calculate mean productive frequencies for all files
top50_clone_freq_mean<-mean(prod_freq_matrix[,"top50_mean"])
top100_clone_freq_mean<-mean(prod_freq_matrix[,"top100_mean"])
top500_clone_freq_mean<-mean(prod_freq_matrix[,"top500_mean"])
all_clone_freq_mean<-mean(prod_freq_matrix[,"all_clones_mean"])

# calculate fraction of clones with freq <0.001
top50_freq_one_tenth<-sum(myfilelist_top50[[1]]$productive_frequency <0.001)/sum(myfilelist_top50[[1]]$productive_frequency>0)
top100_freq_one_tenth<-sum(myfilelist_top100[[1]]$productive_frequency <0.001)/sum(myfilelist_top100[[1]]$productive_frequency>0)
top500_freq_one_tenth<-sum(myfilelist_top500[[1]]$productive_frequency <0.001)/sum(myfilelist_top500[[1]]$productive_frequency>0)
all_freq_one_tenth<-sum(myfilelist[[1]]$productive_frequency <0.001)/sum(myfilelist[[1]]$productive_frequency>0)

# calculate fraction of clones with freq <0.01
top50_freq_one<-sum(myfilelist_top50[[1]]$productive_frequency<0.01)/sum(myfilelist_top50[[1]]$productive_frequency>0)
top100_freq_one<-sum(myfilelist_top100[[1]]$productive_frequency<0.01)/sum(myfilelist_top100[[1]]$productive_frequency>0)
top500_freq_one<-sum(myfilelist_top500[[1]]$productive_frequency<0.01)/sum(myfilelist_top500[[1]]$productive_frequency>0)
all_freq_one<-sum(myfilelist[[1]]$productive_frequency<0.01)/sum(myfilelist[[1]]$productive_frequency>0)


```

```{r}
# make table of original file lengths

total_clones<-matrix(nrow=length(myfilelist),ncol=1)
for (i in 1:length(myfilelist)){
  total_clones[i,1]<-dim(myfilelist[[i]])[[1]]
}

rownames(total_clones)<- NameList
colnames(total_clones)<-c("Total Clones")

#print matrix
print("Total clones in each file")
total_clones %>% 
  DT::datatable()
```

Sept 16,2021
# Make a histogram of the productive frequencies of each sample
```{r}

#create empty data frame
prod_freq_df<-data.frame()


# add productive frequencies to the dataframe
for (i in 1:length(myfilelist)){
  #make a dataframe from the ith productive frequency numbers
  temp_df<-data.frame(productive_frequency=myfilelist[[i]]$productive_frequency)
  #add info about where data came from
  temp_df$fileName <- names(myfilelist[i])
  # merge dataframes to make one large dataframe
  prod_freq_df<- rbind(prod_freq_df,temp_df)
}

ggplot(prod_freq_df, aes(productive_frequency, fill = fileName)) + 
   geom_histogram(alpha = 0.5, bins=100, position = 'identity') +
   scale_x_log10() +
   geom_abline(slope = 1, intercept = 0, linetype=2) +
   theme_classic(base_size = 14) 

ggplot(prod_freq_df, aes(productive_frequency, fill = fileName)) + 
   geom_histogram(alpha = 0.5, bins=100) +
   scale_x_log10() +
   geom_abline(slope = 1, intercept = 0, linetype=2) +
   theme_classic(base_size = 14) 

ggplot(prod_freq_df, aes(productive_frequency, fill = fileName)) + 
   geom_histogram(alpha = 0.5, bins=50, position = 'identity') +
   scale_x_log10() +
   geom_abline(slope = 1, intercept = 0, linetype=2) +
   theme_classic(base_size = 14) 

```

